---
title: "Cybersecurity Assessment Analytics: Drivers of Compliance Scores and Project Delivery Efficiency at Digital Encode Ltd"
author: "Oladeinde Bisola"
date: today
format:
  html:
    theme: flatly
    toc: true
    toc-depth: 3
    code-fold: true
    self-contained: true
execute:
  warning: false
  message: false
---

```{r setup, include=FALSE}
pkgs <- c("tidyverse", "lubridate", "scales", "ggcorrplot",
          "patchwork", "car", "broom", "knitr", "kableExtra", "janitor")
invisible(lapply(pkgs, library, character.only = TRUE))
knitr::opts_chunk$set(fig.width = 9, fig.height = 5, dpi = 150)
```

```{python setup-py, include=FALSE}
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt
import matplotlib.ticker as mticker
import seaborn as sns
from scipy.stats import f_oneway, kruskal, ttest_ind, pearsonr
from statsmodels.stats.multicomp import pairwise_tukeyhsd
import statsmodels.formula.api as smf
import statsmodels.api as sm
from IPython.display import display
import re
sns.set_theme(style="whitegrid", palette="muted")
plt.rcParams["figure.dpi"] = 150
```

```{r load-r, include=FALSE}
# ⚠ Column names are very long with spaces and special characters.
# janitor::clean_names() converts them to clean snake_case automatically.
df_r <- read_csv("data/cybersecurity_assessment report_report_dig_enc_ltd.csv",
                 col_types = cols(.default = col_character())) |>
  clean_names()

# Resulting clean column names (representative):
# timestamp, type_of_cybersecurity_assessment, client_identifier,
# project_delivery_mode, client_industry, client_org_size,
# years_of_assessment, consultants, systems_in_scope,
# meetings, client_responsiveness, security_maturity_level,
# total_duration_of_the_project_days, delays, compliance_score_percent,
# number_of_vulnerabilities_found, number_of_critical_high_vulnerabilities

df_r <- df_r |>
  mutate(
    timestamp                     = dmy_hm(timestamp, quiet=TRUE),
    year                          = year(timestamp),
    years_of_assessment           = as.numeric(years_of_assessment),
    consultants                   = as.numeric(consultants),
    systems_in_scope              = as.numeric(systems_in_scope),
    meetings                      = as.numeric(meetings),
    # Duration and score columns — identified by partial name after clean_names()
    across(contains("duration"),  ~ as.numeric(gsub("[^0-9.]", "", .))),
    across(contains("compliance"), ~ as.numeric(gsub("[^0-9.]", "", .))),
    across(contains("vulnerabilities"), ~ as.numeric(gsub("[^0-9.]", "", .)))
  )

# Identify key columns by partial name
dur_col    <- names(df_r)[str_detect(names(df_r), "duration")][[1]]
comp_col   <- names(df_r)[str_detect(names(df_r), "compliance")][[1]]
vuln_col   <- names(df_r)[str_detect(names(df_r), "^number_of_v")][[1]]
crit_col   <- names(df_r)[str_detect(names(df_r), "critical")][[1]]
assess_col <- names(df_r)[str_detect(names(df_r), "type_of")][[1]]
delivery_col <- names(df_r)[str_detect(names(df_r), "delivery")][[1]]
industry_col <- names(df_r)[str_detect(names(df_r), "industry")][[1]]
resp_col   <- names(df_r)[str_detect(names(df_r), "responsiveness")][[1]]
maturity_col <- names(df_r)[str_detect(names(df_r), "maturity")][[1]]
delays_col <- names(df_r)[str_detect(names(df_r), "^delays")][[1]]
size_col   <- names(df_r)[str_detect(names(df_r), "org_size")][[1]]
scope_col  <- "systems_in_scope"

# Ordered factor for responsiveness and maturity
df_r <- df_r |>
  mutate(
    client_responsiveness_ord = factor(.data[[resp_col]],
      levels = c("Very Low","Low","Moderate","High","Very High"), ordered=TRUE),
    security_maturity_ord = factor(.data[[maturity_col]],
      levels = c("Initial","Developing","Defined","Managed","Optimising"), ordered=TRUE)
  )
```

```{python load-py, include=FALSE}
df_py = pd.read_csv(
    "data/cybersecurity_assessment report_report_dig_enc_ltd.csv",
    dtype=str)

# ⚠ Column names are very long — normalize to snake_case
def to_snake(col):
    col = col.strip().lower()
    col = re.sub(r"[^\w\s]", "", col)     # remove special chars
    col = re.sub(r"\s+", "_", col)         # spaces → underscore
    col = re.sub(r"_+", "_", col).strip("_")
    return col

df_py.columns = [to_snake(c) for c in df_py.columns]

# Parse timestamp
df_py["timestamp"] = pd.to_datetime(df_py["timestamp"].str.strip(),
                                     dayfirst=True, errors="coerce")
df_py["year"] = df_py["timestamp"].dt.year

# Identify key columns by keyword
def fc(keyword, df=df_py):
    return next((c for c in df.columns if keyword in c), None)

dur_col    = fc("duration")
comp_col   = fc("compliance")
vuln_col   = fc("number_of_v")
crit_col   = fc("critical")
assess_col = fc("type_of")
delivery_col = fc("delivery")
industry_col = fc("industry")
resp_col   = fc("responsiveness")
maturity_col = fc("maturity")
delays_col = fc("delays")
size_col   = fc("org_size")
scope_col  = fc("systems_in_scope")
consult_col = fc("consultants")
meetings_col = fc("meetings")

# Convert numeric columns
for col in [dur_col, comp_col, vuln_col, crit_col,
            scope_col, consult_col, meetings_col]:
    if col:
        df_py[col] = pd.to_numeric(
            df_py[col].str.replace(r"[^\d.\-]", "", regex=True), errors="coerce")

resp_order    = ["Very Low","Low","Moderate","High","Very High"]
maturity_order = ["Initial","Developing","Defined","Managed","Optimising"]
```

---

## 1. Executive Summary

> **Instructions:** Complete after your analysis. 150–200 words.

**Business problem:** Digital Encode Ltd's cybersecurity consulting practice delivers assessments across diverse client industries and organisation sizes. Understanding which engagement factors drive compliance score outcomes and project duration efficiency informs staffing decisions, engagement design, and client onboarding processes.

**Data collected:** 100 cybersecurity assessment records collected via standardised Google Form across 2023–2026, covering assessment type, delivery mode, client characteristics, engagement inputs, and outcome metrics including compliance scores and vulnerability counts.

**Key findings:**

1. [Assessment Type X] engagements achieve a mean compliance score of [Y%], compared to [Z%] for [Type W].
2. Client responsiveness shows a [positive] correlation with compliance score (r = [value]).
3. The regression model explains [R²]% of project duration variance, identifying [top driver] as the primary efficiency lever.

**Recommendation:** [Specific operational or engagement design change with projected compliance or duration improvement.]

---

## 2. Professional Disclosure

> **Instructions:** Paste your completed Section 2 from `docs/section_a_30_marks_template.md` here.

---

## 3. Data Collection & Sampling

> **Instructions:** Paste your completed Section 3 from `docs/section_a_30_marks_template.md` here.

---

## 4. Data Description

::: {.panel-tabset}

## R

```{r desc-r}
cat("Rows:", nrow(df_r), "| Columns:", ncol(df_r), "\n\n")

df_r |>
  select(all_of(c(dur_col, comp_col, vuln_col, crit_col,
                  scope_col, "consultants", "meetings"))) |>
  summary() |>
  kable(caption = "Numeric variable summaries") |>
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

```{r desc-cats-r}
df_r |> count(.data[[assess_col]], sort=TRUE) |>
  kable(caption = "Assessment type distribution") |>
  kable_styling(full_width = FALSE)

df_r |> count(.data[[delivery_col]], sort=TRUE) |>
  kable(caption = "Delivery mode distribution") |>
  kable_styling(full_width = FALSE)

df_r |> count(.data[[industry_col]], sort=TRUE) |>
  kable(caption = "Client industry distribution") |>
  kable_styling(full_width = FALSE)
```

## Python

```{python desc-py}
print(f"Rows: {df_py.shape[0]} | Columns: {df_py.shape[1]}\n")
num_cols = [c for c in [dur_col, comp_col, vuln_col, crit_col, scope_col] if c]
display(df_py[num_cols].describe().round(2))
```

```{python desc-cats-py}
for col, label in [(assess_col, "Assessment type"),
                   (delivery_col, "Delivery mode"),
                   (industry_col, "Client industry")]:
    if col:
        print(f"\n{label}:")
        display(df_py[col].value_counts().reset_index(name="count"))
```

:::

---

## 5. Technique 1 — Exploratory Data Analysis

**Business justification:** Profiles the distribution of compliance scores and project durations across the full portfolio before deeper analysis. Identifies whether certain engagement configurations are systematically associated with poor outcomes.

::: {.panel-tabset}

## R

```{r eda-r}
p1 <- ggplot(df_r, aes(x = .data[[comp_col]])) +
  geom_histogram(bins = 20, fill = "#2c7bb6", alpha = 0.8) +
  labs(title = "Distribution of Compliance Score (%)",
       x = "Compliance Score (%)", y = "Count") +
  theme_minimal()

p2 <- ggplot(df_r, aes(x = .data[[dur_col]])) +
  geom_histogram(bins = 20, fill = "#1a9850", alpha = 0.8) +
  labs(title = "Distribution of Project Duration (Days)",
       x = "Total Duration (Days)", y = "Count") +
  theme_minimal()

p1 / p2
```

```{r eda-summary-r}
df_r |>
  group_by(.data[[assess_col]]) |>
  summarise(
    n            = n(),
    mean_score   = mean(.data[[comp_col]], na.rm=TRUE),
    med_score    = median(.data[[comp_col]], na.rm=TRUE),
    mean_dur     = mean(.data[[dur_col]], na.rm=TRUE),
    mean_vuln    = mean(.data[[vuln_col]], na.rm=TRUE),
    .groups = "drop"
  ) |>
  mutate(across(where(is.numeric), ~round(., 1))) |>
  kable(caption = "Compliance & duration summary by assessment type") |>
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

## Python

```{python eda-py}
fig, axes = plt.subplots(1, 2, figsize=(11, 4))
df_py[comp_col].hist(bins=20, ax=axes[0], color="#2c7bb6", edgecolor="white")
axes[0].set_title("Compliance Score Distribution")
axes[0].set_xlabel("Compliance Score (%)")

df_py[dur_col].hist(bins=20, ax=axes[1], color="#1a9850", edgecolor="white")
axes[1].set_title("Project Duration Distribution")
axes[1].set_xlabel("Total Duration (Days)")
plt.tight_layout(); plt.show()
```

```{python eda-summary-py}
summary = df_py.groupby(assess_col)[[comp_col, dur_col, vuln_col]].agg(
    ["mean","median","count"]).round(1)
display(summary)
```

:::

> **Interpretation:** Is the compliance score distribution approximately normal, or skewed? Are there assessments with very low scores (<50%) that warrant attention? Which assessment type has the most variable duration?

---

## 6. Technique 2 — Data Visualization

**Business justification:** Board and client-facing charts showing compliance performance by assessment type, delivery mode, and client responsiveness — directly applicable to practice reviews and client onboarding presentations.

::: {.panel-tabset}

## R

```{r viz-r}
# Box: compliance score by assessment type
p1 <- ggplot(df_r, aes(x = reorder(.data[[assess_col]], .data[[comp_col]], median),
                        y = .data[[comp_col]], fill = .data[[assess_col]])) +
  geom_boxplot(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Compliance Score by Assessment Type",
       x = NULL, y = "Compliance Score (%)") +
  theme_minimal()

# Box: duration by delivery mode
p2 <- ggplot(df_r, aes(x = reorder(.data[[delivery_col]], .data[[dur_col]], median),
                        y = .data[[dur_col]], fill = .data[[delivery_col]])) +
  geom_boxplot(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Project Duration by Delivery Mode",
       x = NULL, y = "Total Duration (Days)") +
  theme_minimal()

p1 / p2
```

```{r viz-maturity-r}
# Bar: mean compliance score by security maturity level
df_r |>
  group_by(.data[[maturity_col]]) |>
  summarise(mean_score = mean(.data[[comp_col]], na.rm=TRUE),
            n = n(), .groups="drop") |>
  ggplot(aes(x = .data[[maturity_col]], y = mean_score, fill = .data[[maturity_col]])) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = sprintf("n=%d\n%.0f%%", n, mean_score)),
            vjust = -0.3, size = 3) +
  labs(title = "Mean Compliance Score by Client Security Maturity Level",
       x = "Security Maturity", y = "Mean Compliance Score (%)") +
  theme_minimal()
```

## Python

```{python viz-py}
fig, axes = plt.subplots(1, 2, figsize=(13, 5))
order1 = df_py.groupby(assess_col)[comp_col].median().sort_values().index
sns.boxplot(data=df_py, y=assess_col, x=comp_col, order=order1, palette="muted", ax=axes[0])
axes[0].set_title("Compliance Score by Assessment Type")
axes[0].set_xlabel("Compliance Score (%)")

order2 = df_py.groupby(delivery_col)[dur_col].median().sort_values().index
sns.boxplot(data=df_py, y=delivery_col, x=dur_col, order=order2, palette="Set2", ax=axes[1])
axes[1].set_title("Project Duration by Delivery Mode")
axes[1].set_xlabel("Total Duration (Days)")
plt.tight_layout(); plt.show()
```

```{python viz-maturity-py}
mat_summary = (df_py.groupby(maturity_col)[comp_col]
               .agg(mean="mean", n="count").reset_index())
# Enforce order if possible
if all(m in mat_summary[maturity_col].values for m in maturity_order[:1]):
    mat_summary[maturity_col] = pd.Categorical(
        mat_summary[maturity_col], categories=maturity_order, ordered=True)
    mat_summary = mat_summary.sort_values(maturity_col)

fig, ax = plt.subplots(figsize=(9, 4))
ax.bar(range(len(mat_summary)), mat_summary["mean"], color=sns.color_palette("muted"))
ax.set_xticks(range(len(mat_summary)))
ax.set_xticklabels(mat_summary[maturity_col], rotation=20, ha="right")
ax.set_title("Mean Compliance Score by Security Maturity Level")
ax.set_ylabel("Mean Compliance Score (%)")
plt.tight_layout(); plt.show()
```

:::

> **Interpretation:** Is there a visible step-up in compliance score as client security maturity increases? Do remote-delivered engagements take longer than on-site? Identify any assessment type where the median and mean diverge significantly.

---

## 7. Technique 3 — Hypothesis Testing

**Business justification:** Formally tests whether compliance score differences across assessment types and delivery modes are statistically meaningful, providing the evidential basis for service design recommendations.

::: {.panel-tabset}

## R

```{r hyp-r}
# ANOVA: compliance score by assessment type
aov_at <- aov(df_r[[comp_col]] ~ df_r[[assess_col]])
cat("=== ANOVA: Compliance Score by Assessment Type ===\n")
summary(aov_at)

eta_sq <- tidy(aov_at)$sumsq[1] / sum(tidy(aov_at)$sumsq)
cat(sprintf("\nEta-squared: %.4f (%s effect)\n", eta_sq,
    ifelse(eta_sq < 0.01, "negligible",
    ifelse(eta_sq < 0.06, "small",
    ifelse(eta_sq < 0.14, "medium", "large")))))

tukey_at <- TukeyHSD(aov_at)
tidy(tukey_at) |>
  arrange(adj.p.value) |>
  kable(digits=4, caption="Tukey HSD: Compliance Score by Assessment Type") |>
  kable_styling(bootstrap_options="striped", full_width=FALSE)
```

```{r hyp-mode-r}
# ANOVA: project duration by delivery mode
aov_dm <- aov(df_r[[dur_col]] ~ df_r[[delivery_col]])
cat("=== ANOVA: Project Duration by Delivery Mode ===\n")
summary(aov_dm)

# Kruskal-Wallis robustness check
kw <- kruskal.test(df_r[[comp_col]] ~ df_r[[assess_col]])
cat(sprintf("\nKruskal-Wallis (compliance ~ assessment type): chi-sq=%.4f, p=%.4f\n",
            kw$statistic, kw$p.value))
```

## Python

```{python hyp-py}
at_groups = [g[comp_col].dropna().values
             for _, g in df_py.groupby(assess_col)]
f_stat, p_val = f_oneway(*at_groups)
print(f"=== ANOVA: Compliance Score by Assessment Type ===")
print(f"F = {f_stat:.4f}, p = {p_val:.4f}\n")

tukey = pairwise_tukeyhsd(
    endog=df_py[comp_col].dropna(),
    groups=df_py.loc[df_py[comp_col].notna(), assess_col]
)
print(tukey.summary())
```

```{python hyp-dur-py}
dm_groups = [g[dur_col].dropna().values
             for _, g in df_py.groupby(delivery_col)]
f2, p2 = f_oneway(*dm_groups)
print(f"\nANOVA: Duration by Delivery Mode: F={f2:.4f}, p={p2:.4f}")

h_stat, p_kw = kruskal(*at_groups)
print(f"Kruskal-Wallis (compliance ~ assessment type): H={h_stat:.4f}, p={p_kw:.4f}")
```

:::

> **Interpretation:** State H₀ and H₁. Report the p-value and effect size. Write one business sentence per significant finding. Example: "Penetration testing engagements produce significantly lower compliance scores than compliance audits (p = 0.02), consistent with their more adversarial testing methodology — this should be disclosed in client engagement letters."

---

## 8. Technique 4 — Correlation Analysis

**Business justification:** Identifies which engagement inputs (systems in scope, consultant count, meetings, client responsiveness) have the strongest relationships with compliance score and project duration outcomes.

::: {.panel-tabset}

## R

```{r corr-r}
num_cols_r <- df_r |>
  select(all_of(c(comp_col, dur_col, vuln_col, crit_col,
                  scope_col, "consultants", "meetings"))) |>
  drop_na()

cor_mat <- cor(num_cols_r, method = "pearson")

ggcorrplot(cor_mat, method = "circle", type = "lower",
           lab = TRUE, lab_size = 3.5,
           colors = c("#d73027", "white", "#1a9850"),
           title = "Pearson Correlation — Cybersecurity Assessment Variables",
           ggtheme = theme_minimal(base_size = 10))
```

```{r corr-table-r}
as.data.frame(cor_mat) |>
  rownames_to_column("Variable") |>
  select(Variable, all_of(comp_col)) |>
  filter(Variable != comp_col) |>
  arrange(desc(abs(.data[[comp_col]]))) |>
  rename(r_with_compliance_score = .data[[comp_col]]) |>
  kable(digits=4, caption="Correlation with Compliance Score (%)") |>
  kable_styling(full_width=FALSE)
```

## Python

```{python corr-py}
num_cols_py = [c for c in [comp_col, dur_col, vuln_col, crit_col,
                             scope_col, consult_col, meetings_col] if c]
corr_data = df_py[num_cols_py].dropna()
corr_mat  = corr_data.corr(method="pearson")

fig, ax = plt.subplots(figsize=(8, 6))
mask = np.triu(np.ones_like(corr_mat, dtype=bool))
sns.heatmap(corr_mat, mask=mask, annot=True, fmt=".2f",
            cmap="RdYlGn", center=0, linewidths=0.5, ax=ax,
            annot_kws={"size": 8})
ax.set_title("Pearson Correlation — Assessment Variables")
plt.tight_layout(); plt.show()

print(f"\nCorrelation with compliance score ({comp_col}):")
print(corr_mat[comp_col].drop(comp_col).sort_values(key=abs, ascending=False).round(4))
```

:::

> **Interpretation:** Does the number of systems in scope correlate with project duration? Does client responsiveness (if encoded numerically) correlate with compliance score? Flag any unexpected correlations between input variables that may indicate multicollinearity.

---

## 9. Technique 5 — Multiple Regression

**Business justification:** Produces an engagement efficiency model — given assessment type, delivery mode, client responsiveness, scope, and maturity, the model estimates expected project duration and compliance score, enabling staffing and scheduling decisions.

::: {.panel-tabset}

## R

```{r reg-r}
reg_data <- df_r |>
  mutate(
    assess_f   = factor(.data[[assess_col]]),
    delivery_f = factor(.data[[delivery_col]]),
    maturity_f = factor(.data[[maturity_col]],
                        levels = c("Initial","Developing","Defined","Managed","Optimising"))
  ) |>
  select(all_of(c(dur_col, comp_col, scope_col, "consultants", "meetings")),
         assess_f, delivery_f, maturity_f) |>
  drop_na()

# Model 1: Project duration
model_dur <- lm(reformulate(
  c("assess_f", "delivery_f", scope_col, "consultants", "meetings"),
  response = dur_col), data = reg_data)

cat("=== Regression: Total Project Duration (Days) ===\n")
summary(model_dur)
```

```{r reg-duration-table-r}
tidy(model_dur, conf.int=TRUE) |>
  mutate(across(where(is.numeric), ~round(., 3))) |>
  kable(caption="Regression: Project Duration — coefficients") |>
  kable_styling(bootstrap_options="striped", full_width=FALSE)

glance(model_dur) |>
  select(r.squared, adj.r.squared, p.value, AIC) |>
  mutate(across(everything(), ~round(., 4))) |>
  kable(caption="Model fit: Project Duration") |>
  kable_styling(full_width=FALSE)
```

```{r reg-compliance-r}
# Model 2: Compliance score
model_comp <- lm(reformulate(
  c("assess_f", "delivery_f", "maturity_f", "meetings"),
  response = comp_col), data = reg_data)

cat("\n=== Regression: Compliance Score (%) ===\n")
summary(model_comp)

tidy(model_comp, conf.int=TRUE) |>
  mutate(across(where(is.numeric), ~round(., 3))) |>
  kable(caption="Regression: Compliance Score — coefficients") |>
  kable_styling(bootstrap_options="striped", full_width=FALSE)

glance(model_comp) |>
  select(r.squared, adj.r.squared, p.value, AIC) |>
  mutate(across(everything(), ~round(., 4))) |>
  kable(caption="Model fit: Compliance Score") |>
  kable_styling(full_width=FALSE)
```

```{r reg-diag-r}
par(mfrow = c(2, 2))
plot(model_dur, which = 1:4, pch = 16, cex = 0.6,
     main = "Diagnostics: Project Duration Model")
par(mfrow = c(1, 1))
```

## Python

```{python reg-py}
reg_py = df_py[[dur_col, comp_col, assess_col, delivery_col,
                maturity_col, scope_col, consult_col, meetings_col]].dropna()

formula_dur = (f"{dur_col} ~ C({assess_col}) + C({delivery_col}) + "
               f"{scope_col} + {consult_col} + {meetings_col}")
model_dur_py = smf.ols(formula_dur, data=reg_py).fit()
print("=== Regression: Project Duration ===")
print(model_dur_py.summary())
```

```{python reg-compliance-py}
formula_comp = (f"{comp_col} ~ C({assess_col}) + C({delivery_col}) + "
                f"C({maturity_col}) + {meetings_col}")
model_comp_py = smf.ols(formula_comp, data=reg_py).fit()
print("\n=== Regression: Compliance Score ===")
print(model_comp_py.summary())
```

```{python reg-diag-py}
fig, axes = plt.subplots(1, 2, figsize=(12, 4))
axes[0].scatter(model_dur_py.fittedvalues, model_dur_py.resid, alpha=0.5, s=20)
axes[0].axhline(0, color="red", linestyle="--")
axes[0].set_title("Residuals vs Fitted — Duration Model")
axes[0].set_xlabel("Fitted"); axes[0].set_ylabel("Residuals")

sm.qqplot(model_dur_py.resid, line="s", ax=axes[1], alpha=0.5)
axes[1].set_title("Q-Q Plot — Duration Model")
plt.tight_layout(); plt.show()
print(f"Duration model — Adj. R²: {model_dur_py.rsquared_adj:.4f}")
print(f"Compliance model — Adj. R²: {model_comp_py.rsquared_adj:.4f}")
```

:::

> **Interpretation:** For each significant coefficient, write one business sentence. Example: "Each additional system in scope adds approximately [X] days to project duration (β = X, p < 0.05), suggesting that engagements with more than [N] systems should be scoped for a larger consultant team at proposal stage."

---

## 10. Integrated Findings

> Write 150–200 words connecting the five analyses into one coherent narrative and a single prioritized recommendation.

---

## 11. Limitations & Further Work

- The Google Form survey approach ensures standardised input but relies on self-reported data from project consultants, who may under-report delays or over-report compliance scores. An independent validation of a random subsample would strengthen reliability.
- 100 records across four assessment types and three delivery modes means some subgroups have small sample sizes; the ANOVA results should be interpreted cautiously for groups with n < 15.
- The client responsiveness variable is ordinal — treating it as continuous in correlation/regression assumes equal intervals between levels. An ordinal logistic regression would be a more rigorous alternative.
- Future work: link assessment records to client retention or remediation follow-up data to test whether higher compliance scores predict re-engagement, enabling a revenue forecasting model.

---

## References

> *(Complete in APA format)*

---

## Appendix: AI Usage Statement

> *(One paragraph describing AI assistance and independent judgement.)*
